---
title: "scRNA glioma pipeline"
author: "Stefano de Pretis"
date: "2024-02-15"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(results = FALSE, message=FALSE, warning=FALSE)
```

```{r}
source('SourceFunctions.R')
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggplotify))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(patchwork))
```

# Load data

```{r}
folders <- c("M2_veh","M16_veh","M3","M4","M2","M12","M5","M6")
```


```{r}
samples <- lapply(folders,
       function(project) CreateSeuratObject(
         Read10X(paste0('InputData/',project,'/outs/filtered_feature_bc_matrix/')), 
         project = project, min.cells = 3, min.features = 200
         ))
names(samples) <- folders
```

# QC and selecting cells for further analysis

```{r, fig.height = 6, fig.width = 14, echo = FALSE}
for( s in 1:length(samples) )
  samples[[s]][["percent.mt"]] <- PercentageFeatureSet(samples[[s]], pattern = "^mt-")

scat_count_mt <- lapply(1:length(samples), function(i) {
  s <- samples[[i]]
  s$filter <- c('discarded','selected')[1+1*(s$nFeature_RNA>700)]
  ggplot(s@meta.data, aes(nFeature_RNA, percent.mt, 
                          color = filter
                          )) + 
    geom_point(size=.2) + # scale_color_viridis() +
    ggtitle(paste0(names(samples)[i], ' (', length(which(s$nFeature_RNA>700)),' selected cells)')) + 
    geom_vline(xintercept=700, #linetype="dashed", 
               color = "red") + 
    theme_bw() # + theme(legend.position = 'none')
})

p <- scat_count_mt[[1]]
for(s in 2:length(samples)) p <- p + (scat_count_mt[[s]])
p + plot_layout(nrow=2, guides = 'collect')

ggsave('Figures/scatter_nFeature_percent_mt.png', height = 6, width = 14)
```

```{r, eval=FALSE}
samples_subset <- lapply(samples, subset, subset = nFeature_RNA > 700)
saveRDS(samples_subset, file='Data/samples_features_filtered.rds')
```

```{r}
##  analyze single samples 
# samples_subset <- lapply(samples_subset, function(x) {
#   x <- NormalizeData(x)
#   x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 4000)
#   x <- ScaleData(x, features = rownames(x))
#   x <- RunPCA(x, npcs = 30)
#   x <- RunUMAP(x, reduction = "pca", dims = 1:20)
#   x <- FindNeighbors(x, reduction = "pca", dims = 1:20)
#   x <- FindClusters(x, resolution = .8)
#   return(x)
# })
```

# Batch removal

```{r, eval=FALSE}
samples_subset <- readRDS(file='Data/samples_features_filtered.rds')
samples <- lapply(samples, function(x) {
  # x <- NormalizeData(x) ALREADY NORMALIZED
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 4000)
})
nDims <- 20
features <- FindIntegrationAnchors(object.list = samples,
                                   dims = 1:nDims, anchor.features = 4000)
features.to.integrate = features@anchor.features
glioma_integrated <- IntegrateData(anchorset = features, dims = 1:nDims,
                                   features.to.integrate = features.to.integrate)
DefaultAssay(glioma_integrated) <- "integrated"
saveRDS(glioma_integrated, file='Data/glioma_features_integrated.rds')
```

# Standard workflow for visualization and clustering

```{r, eval=FALSE}
glioma_integrated <- readRDS(file='Data/glioma_features_integrated.rds')
glioma_integrated <- ScaleData(glioma_integrated, features = rownames(glioma_integrated))
glioma_integrated <- RunPCA(glioma_integrated, npcs = 30)
glioma_integrated <- RunUMAP(glioma_integrated, reduction = "pca", dims = 1:20)
glioma_integrated <- FindNeighbors(glioma_integrated, reduction = "pca", dims = 1:20)
glioma_integrated <- FindClusters(glioma_integrated, resolution = .8)
glioma_integrated$orig.ident <- factor(glioma_integrated$orig.ident, levels = c("M2_veh","M16_veh","M3","M4","M2","M12","M5","M6"))
saveRDS(glioma_integrated, file='Data/glioma_features_integrated_clustered.rds')
```

# Assignment of cell-types associated with CelliD using PanglaoDB_markers

```{r, eval=FALSE}
glioma_integrated <- readRDS(
  file='Data/glioma_features_integrated_clustered.rds')
# ---------------
library(CelliD)
glioma_integrated <- RunMCA(glioma_integrated)
panglao <- read_tsv("https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz")
panglao_all <- panglao %>%  filter(str_detect(species,"Mm"))
panglao_all$`official gene symbol` <- str_to_title(panglao_all$`official gene symbol`)
panglao_all <- panglao_all %>% group_by(`cell type`) %>% summarise(geneset = list(`official gene symbol`))
all_gs <- setNames(panglao_all$geneset, panglao_all$`cell type`)
HGT_panglao_all <- RunCellHGT(glioma_integrated, pathways = all_gs, dims = 1:50)
HGT_pangall_prediction <- rownames(HGT_panglao_all)[apply(HGT_panglao_all, 2, which.max)]
HGT_pangall_prediction <- ifelse(apply(HGT_panglao_all, 2, max)>2, yes = HGT_pangall_prediction, "unassigned")
HGT_pangall_prediction[HGT_pangall_prediction == 'Monocytes'] <- "unassigned"
glioma_integrated$cellid_panall <- filterLowFreq(HGT_pangall_prediction, 50)
glioma_integrated$cellid_panall_conf <- apply(HGT_panglao_all, 2, max)
# Define color palette
levels <- sort(unique(glioma_integrated$cellid_panall))
cols_panall <- ggplotColours(length(levels))
cols_panall[levels=='unassigned'] <- 'grey60'
cols_panall[levels=='Neurons'] <- 'gold'
cols_panall[levels=='Macrophages'] <- 'deepskyblue4'
Misc(glioma_integrated, 'cols_panall') <- cols_panall
# ---------
saveRDS(glioma_integrated, file='Data/glioma_features_integrated_clustered_cellid.rds')
```

# Import Souporcell genotypes 

```{r}
glioma_integrated <- readRDS(file='Data/glioma_features_integrated_clustered_cellid.rds')
# -----------
sorted_samples <- levels(glioma_integrated$orig.ident)
filenames <- paste0('InputData/', sorted_samples, '/outs/soup_or_cell/clusters.tsv')
souporcell_all <- lapply(filenames, function(file) {
  x <- read.table(file, header = T, row.names = 1)
  tumor_class <- names(which.min(table(x$assignment)[c('0','1')]))
  x$tumor_assignment <- ifelse(x$assignment == tumor_class, 'tumor', 'normal')
  return(x)
  })
names(souporcell_all) <- sorted_samples
#------
for(i in seq(souporcell_all)) rownames(souporcell_all[[i]]) <- 
  paste(rownames(souporcell_all[[i]]), i, sep = '_')
names(souporcell_all) <- NULL
souporcell_merged <- do.call('rbind', souporcell_all)
glioma_integrated <- AddMetaData(glioma_integrated, souporcell_merged)
glioma_integrated$tumor_assignment[glioma_integrated$status != 'singlet'] <- NA

# assign the likelihood of being tumor/normal

meta_data <- glioma_integrated@meta.data
meta_data$tumor_likelihood <- meta_data$normal_likelihood <- NA
meta_data$normal_likelihood <- meta_data$normal_likelihood <- NA
meta_data_new <- do.call('rbind', lapply(unique(glioma_integrated$orig.ident), function(s) {
  sample_meta <- subset(meta_data, orig.ident==s)
  df <- unique(data.frame(t=sample_meta$tumor_assignment, a=sample_meta$assignment))
  df <- na.omit(df)
  sample_meta[[paste(df[1,1], 'likelihood', sep='_')]] <- sample_meta[[paste0('cluster', df[1,2])]]
  sample_meta[[paste(df[2,1], 'likelihood', sep='_')]] <- sample_meta[[paste0('cluster', df[2,2])]]
  return(sample_meta)
}))
meta_data_new <- meta_data_new[rownames(meta_data),]
glioma_integrated$tumor_likelihood <- meta_data_new$tumor_likelihood
glioma_integrated$normal_likelihood <- meta_data_new$normal_likelihood
glioma_integrated$tumor_assignment[is.na(glioma_integrated$tumor_assignment)] <- 'mixed'
glioma_integrated$genotype <- glioma_integrated$tumor_assignment
```

```{r, fig.height=6, fig.width=10}
orig_idents <- levels(glioma_integrated$orig.ident)
gg_lik <- ggplot(subset(glioma_integrated@meta.data, orig.ident == orig_idents[1]), 
       aes(tumor_likelihood, normal_likelihood, color=genotype)) + 
  geom_point() + geom_abline(slope = 1, linetype='dashed', linewidth=.2) + theme_bw() + scale_color_manual( values = c('mixed' = 'gold', normal='grey80',tumor='grey30')) + ggtitle(orig_idents[1])

for(i in 2:length(orig_idents)) {
  gg_lik_i <- ggplot(subset(glioma_integrated@meta.data, orig.ident == orig_idents[i]), 
         aes(tumor_likelihood, normal_likelihood, color=genotype)) + 
    geom_point() + geom_abline(slope = 1, linetype='dashed', linewidth=.2) + theme_bw() + scale_color_manual( values = c('mixed' = 'gold', normal='grey80',tumor='grey30')) + ggtitle(orig_idents[i])
  gg_lik <- gg_lik + gg_lik_i
}
gg_lik + plot_layout(nrow=2, guides = 'collect')

ggsave('Figures/scatter_nFeature_percent_mt.png', height = 6, width = 15)
```


```{r, fig.height=3, fig.width=10}
p1 <- DimPlot(glioma_integrated, group.by = 'tumor_assignment') + 
  scale_color_manual(values = c('mixed' = 'gold', normal='grey80',tumor='grey30'))

p2 <- FeaturePlot(glioma_integrated, 'nFeature_RNA')

p3 <- FeaturePlot(glioma_integrated, 'percent.mt')

p1 + p2 + p3 + plot_layout(nrow = 1)

ggsave('Figures/umap_integrated_clustered.png', height = 3, width = 10)
```

```{r, eval=FALSE}
saveRDS(glioma_integrated, 
file='Data/glioma_features_integrated_clustered_cellid_genotyped.rds')
```

# remove cells that cluster with the other genotype

```{r, message=TRUE}
glioma_integrated <- readRDS(
  file='RerunData/glioma_features_integrated_clustered_cellid_genotyped.rds')
t_n_clust <- table(glioma_integrated$tumor_assignment, glioma_integrated$seurat_clusters)
# identify tumor and normal clusters
t_n_groups <- split(colnames(t_n_clust), rownames(t_n_clust)[apply(t_n_clust,2,which.max)])
# remove cells that do not map into the cluster of their class
normal_cluster <- glioma_integrated$seurat_clusters %in% t_n_groups$normal
normal_cell <- glioma_integrated$tumor_assignment == 'normal'
normal_cell[glioma_integrated$tumor_assignment == 'mixed'] <- NA
coherent_cells <- normal_cluster == normal_cell
message('percentage of mixed genotypes cells:               ', round(100*(length(which(is.na(coherent_cells)))/length(coherent_cells)),1), '%')
message('percentage of cells with incoherent clustering:    ', round(100*(length(which(!coherent_cells))/length(which(!is.na(coherent_cells)))),1), '%')
```


```{r, eval=FALSE}
# Run the standard workflow for visualization and clustering
glioma_integrated_filtered <- glioma_integrated[,coherent_cells]
glioma_integrated_filtered <- ScaleData(glioma_integrated_filtered, features = rownames(glioma_integrated_filtered))
glioma_integrated_filtered <- RunPCA(glioma_integrated_filtered, npcs = 30)
glioma_integrated_filtered <- RunUMAP(glioma_integrated_filtered, reduction = "pca", dims = 1:20)
glioma_integrated_filtered <- FindNeighbors(glioma_integrated_filtered, reduction = "pca", dims = 1:20)
glioma_integrated_filtered <- FindClusters(glioma_integrated_filtered, resolution = .8)
# -------------
saveRDS(glioma_integrated_filtered, file='Data/glioma_features_integrated_clustered_cellid_genotyped_filtered.rds')
```

# report

```{r}
# load the dataset
glioma_integrated <- readRDS(file='Data/glioma_features_integrated_clustered_cellid_genotyped_filtered.rds')
# collapse the replicates into their treatment
glioma_integrated$treatment <- 'untreated'
glioma_integrated$treatment[glioma_integrated$orig.ident %in% c('M3', 'M4')] <- 'RT'
glioma_integrated$treatment[glioma_integrated$orig.ident %in% c('M2', 'M12')] <- 'RT+TMZ'
glioma_integrated$treatment[glioma_integrated$orig.ident %in% c('M5', 'M6')] <- 'RT+TMZ+MET'
glioma_integrated$treatment <- factor(glioma_integrated$treatment, levels = c('untreated','RT','RT+TMZ','RT+TMZ+MET'))
```

```{r}
# # make a palette for treatment
# treatment_palette <- brewer.pal(4, 'Spectral')
# names(treatment_palette) <- levels(glioma_integrated$treatment)
# glioma_integrated@misc$cols_treatment <- treatment_palette
```


```{r}
# assign the likelihood of being tumor/normal
meta_data <- glioma_integrated@meta.data
meta_data$tumor_likelihood <- meta_data$normal_likelihood <- NA
meta_data$normal_likelihood <- meta_data$normal_likelihood <- NA
meta_data_new <- do.call('rbind', lapply(unique(glioma_integrated$orig.ident), function(s) {
  sample_meta <- subset(meta_data, orig.ident==s)
  df <- unique(data.frame(t=sample_meta$tumor_assignment, a=sample_meta$assignment))
  sample_meta[[paste(df[1,1], 'likelihood', sep='_')]] <- sample_meta[[paste0('cluster', df[1,2])]]
  sample_meta[[paste(df[2,1], 'likelihood', sep='_')]] <- sample_meta[[paste0('cluster', df[2,2])]]
  return(sample_meta)
}))
meta_data_new <- meta_data_new[rownames(meta_data),]
glioma_integrated$tumor_likelihood <- meta_data_new$tumor_likelihood
glioma_integrated$normal_likelihood <- meta_data_new$normal_likelihood
# assign first level to tumor (for ggplot coloring reasons)
glioma_integrated$tumor_assignment <- factor(glioma_integrated$tumor_assignment, levels = c('tumor', 'normal'))
```

```{r}
# define cluster colors
cols_seurat <- c(palette[1:24], palette[1:6])
names(cols_seurat) <- levels(glioma_integrated$seurat_clusters)
glioma_integrated@misc$cols_seurat <- cols_seurat
```

# Genotype analysis

## UMAP

```{r}
DimPlot(glioma_integrated, group.by = 'tumor_assignment') +
  scale_color_manual(values = c(normal='grey80',tumor='grey30'))
ggsave('Figures/umap_filtered_tumor.png', height = 4, width = 4.5)
```

## likelihood

```{r}
ggplot(glioma_integrated@meta.data, aes(tumor_likelihood, normal_likelihood, color=tumor_assignment)) + 
  geom_point() + geom_abline(slope = 1, linetype='dashed', size=.2) + theme_bw() + scale_color_manual(values = c(normal='grey80',tumor='grey30'))
ggsave('Figures/tumor_likelihood_filtered_cells.png')
```

## UMAP by treatment

```{r}
DimPlot(glioma_integrated, group.by = 'tumor_assignment', split.by = 'treatment', ncol=2) + scale_color_manual(values = c(normal='grey80',tumor='grey30'))
ggsave('Figures/umap_filtered_tumor_splitTreatment.png', height = 7, width = 7)
```

## barplot


```{r}
glioma_integrated$tumor_assignment <- 
  factor(glioma_integrated$tumor_assignment, levels = c('tumor', 'normal'))
treat_tum_table <- table(glioma_integrated$treatment, glioma_integrated$tumor_assignment)
treat_tum_fract <- t(treat_tum_table/rowSums(treat_tum_table))
bpOut <- barplot(treat_tum_fract, col = c('grey30','grey80'), 
                 legend.text = c('tumor', 'normal'), ylim=c(0,1.1), args.legend = list(x='right'))
text(bpOut, rep(1, length(bpOut)), paste(round(100*treat_tum_fract[1,],1), '%'), pos = 3)
```

# Cell type identification

## UMAP - Panglao database

```{r, fig.height=5, fig.width=11}
DimPlot(glioma_integrated, group.by = 'cellid_panall', cols = glioma_integrated@misc$cols_panall) + 
  FeaturePlot(glioma_integrated, features = 'cellid_panall_conf') + 
  plot_layout(guides = 'collect')
ggsave('Figures/umap_filtered_panglao.png', height = 6, width = 12)
```

# Clustering

## UMAP

```{r}
DimPlot(glioma_integrated, label = T, cols = glioma_integrated@misc$cols_seurat)
ggsave('Figures/umap_filtered_clusters.png', height = 4.5, width = 6)
```

## Divide clusters into groups

```{r}
tumor_group <- c('5','8','9','10')
lymphatic_group <- c('1','4','6','16','21','22')
parenchymal_group <- c('11','13','17','23','25','28','29')
myeloid <- setdiff(as.character(sort(unique(glioma_integrated$seurat_clusters))),
                   c(tumor_group, lymphatic_group, parenchymal_group))
Microglia_group <-  c('0', '2', '7', '3', '12', '15', '18', '20')
DCs_group  <- c('14', '19', '24', '26', '27')
all_groups <- c(tumor_group, lymphatic_group, parenchymal_group, 
  Microglia_group, DCs_group)
# check that cluster are represented just once
stopifnot(all(!duplicated(all_groups)))
# check that all cluster are represented within groups
stopifnot(length(setdiff(as.character(sort(unique(glioma_integrated$seurat_clusters))), 
        all_groups)) == 0)
```


```{r}
cluster_groups <- list(tumor=tumor_group, 
                       lymphatic=lymphatic_group, 
                       parenchymal=parenchymal_group, 
                       myeloid=myeloid)
                       # MoTAM=MoTAM_group,
                       # Microglia=Microglia_group,
                       # DCs=DCs_group)
                       
cell_groups <- lapply(cluster_groups, function(k) {
  colnames(glioma_integrated)[glioma_integrated$seurat_clusters %in% k]
}); names(cell_groups) <- names(cluster_groups)

glioma_integrated@misc$cols_groups <- c(
  'tumor'='grey30',
  'lymphatic'='#19b9c2',
  'parenchymal'='#c2a019',
  'myeloid'='#ceb4db'
  # 'MoTAM'='#ceb4db',
  # 'Microglia'='#ceb4db',
  # 'DCs'='#ceb4db'
)

group1 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[1], 
                  cols.highlight = 'grey30') + NoLegend() + ggtitle('tumor group')
group2 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[2], 
                  cols.highlight = '#19b9c2') + NoLegend() + ggtitle('lymphatic group')
group3 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[3], 
                  cols.highlight = '#c2a019') + NoLegend() + ggtitle('parenchymal group') 
group4 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[4],
                  cols.highlight = '#ceb4db') + NoLegend() + ggtitle('myeloid group')
# group4 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[4], 
#                   cols.highlight = '#ceb4db') + NoLegend() + ggtitle('MoTAM')
# group5 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[5], 
#                   cols.highlight = '#ceb4db') + NoLegend() + ggtitle('Microglia')
# group6 <- DimPlot(glioma_integrated, cells.highlight = cell_groups[6], 
#                   cols.highlight = '#ceb4db') + NoLegend() + ggtitle('DCs')

group1 + group2 + group3 + group4 # + group5 + group6

ggsave('Figures/umap_filtered_groups.png', height = 7, width = 8)
```


```{r, eval=FALSE}
# Cluster markers
All_marker = FindAllMarkers(integrated, assay = 'RNA')
write.csv(All_marker, "Tables/Cluster_markers.csv")
```


```{r}
cl_markers <- read.csv('Tables/Cluster_markers.csv', row.names = 1)
cl_markers <- cl_markers[grep('Hb[ab]-',cl_markers$gene, invert = TRUE),]
top10_cl_markers <- subset(cl_markers, !grepl('^mt-', rownames(cl_markers))) %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
```

```{r}
glioma_integrated_RNA <- glioma_integrated
DefaultAssay(glioma_integrated_RNA) <- 'RNA'
glioma_integrated_RNA <- ScaleData(glioma_integrated_RNA, features = rownames(glioma_integrated_RNA))
```

### Heatmap - tumor group

```{r}
top10_cl_markers_tumor <- subset(top10_cl_markers, cluster %in% tumor_group)$gene
glioma_tumor_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% tumor_group]
```

```{r, fig.height=6, fig.width=6}
markers_tumor <- 
  DoAverageHeatmap(object = glioma_tumor_subset, 
                 features = top10_cl_markers_tumor,  
                 cluster_rows = F, cluster_cols = F,
                 gaps_row = seq(0,length(top10_cl_markers_tumor),by=10),
                 gaps_col = seq(unique(glioma_tumor_subset$seurat_clusters)),
                 cluster_palette = glioma_integrated@misc$cols_seurat)
ggsave('Figures/markers_tumor.png', 
       ggplotify::as.ggplot(markers_tumor$gtable),
       height = 10, width = 7)
```

### Heatmap - lymphatic group

```{r, fig.height=13, fig.width=8}
top10_cl_markers_lymphatic <- subset(top10_cl_markers, cluster %in% lymphatic_group)$gene
glioma_lymphatic_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% lymphatic_group]
```

```{r, fig.height=11, fig.width=6}
markers_lymphatic <- 
  DoAverageHeatmap(object = glioma_lymphatic_subset, 
                 features = top10_cl_markers_lymphatic,  
                 cluster_rows = F, cluster_cols = F,
                 gaps_row = seq(0,length(top10_cl_markers_lymphatic),by=10),
                 gaps_col = seq(unique(glioma_lymphatic_subset$seurat_clusters)),
                 cluster_palette = glioma_integrated@misc$cols_seurat)
ggsave('Figures/markers_lymphatic.png', 
       ggplotify::as.ggplot(markers_lymphatic$gtable),
       height = 10, width = 7)
```

### Heatmap - parenchymal group

```{r, fig.height=8, fig.width=6}
top10_cl_markers_parenchymal <- subset(top10_cl_markers, cluster %in% parenchymal_group)$gene
glioma_parenchymal_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% parenchymal_group]
```

```{r, fig.height=11, fig.width=6}
markers_parenchyma <- 
  DoAverageHeatmap(object = glioma_parenchymal_subset, 
                 features = top10_cl_markers_parenchymal,  
                 cluster_rows = F, cluster_cols = F,
                 gaps_row = seq(0,length(top10_cl_markers_parenchymal),by=10),
                 gaps_col = seq(unique(glioma_parenchymal_subset$seurat_clusters)),
                 cluster_palette = glioma_integrated@misc$cols_seurat)
ggsave('Figures/markers_parenchyma.png', 
       ggplotify::as.ggplot(markers_parenchyma$gtable),
       height = 10, width = 7)
```

### Heatmap - myeloid group (Microglia)

```{r, fig.height=11, fig.width=8}
top10_cl_markers_Microglia<- subset(top10_cl_markers, cluster %in% Microglia_group)$gene
glioma_Microglia_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% Microglia_group]
```

```{r, fig.height=11, fig.width=6}
markers_microglia <- 
  DoAverageHeatmap(object = glioma_Microglia_subset, 
                 features = top10_cl_markers_Microglia,  
                 cluster_rows = F, cluster_cols = F,
                 gaps_row = seq(0,length(top10_cl_markers_Microglia),by=10),
                 gaps_col = seq(unique(glioma_Microglia_subset$seurat_clusters)),
                 cluster_palette = glioma_integrated@misc$cols_seurat)
ggsave('Figures/markers_microglia.png', 
       ggplotify::as.ggplot(markers_microglia$gtable),
       height = 12, width = 7)
```

### Heatmap - myeloid group (DCs)

```{r, fig.height=11, fig.width=8}
top10_cl_markers_DCs<- subset(top10_cl_markers, cluster %in% DCs_group)$gene
glioma_DCs_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% DCs_group]
```

```{r, fig.height=11, fig.width=6}
markers_DCs <- 
  DoAverageHeatmap(object = glioma_DCs_subset, 
                 features = top10_cl_markers_DCs,  
                 cluster_rows = F, cluster_cols = F,
                 gaps_row = seq(0,length(top10_cl_markers_DCs),by=10),
                 gaps_col = seq(unique(glioma_DCs_subset$seurat_clusters)),
                 cluster_palette = glioma_integrated@misc$cols_seurat)
ggsave('Figures/markers_DCs.png', 
       ggplotify::as.ggplot(markers_DCs$gtable),
       height = 12, width = 7)
```

## Barplots

### Per group, relative Barplots

```{r, fig.height=7, fig.width=10, message=FALSE}
library(ggplot2)
library(ggalluvial)

glioma_myeloid_subset <- glioma_integrated_RNA[,
         glioma_integrated_RNA$seurat_clusters %in% myeloid]

glioma_subsets <- list(
  tumor=glioma_tumor_subset, 
  lymphatic=glioma_lymphatic_subset, 
  parenchymal=glioma_parenchymal_subset, 
  myeloid=glioma_myeloid_subset)

relative_plots <- function(seuratObj) {
  pt <- table(seuratObj$treatment, droplevels(seuratObj$seurat_clusters))
  pt <- t(as.matrix(pt)/rowSums(pt))
  pt_df <- as.data.frame(pt)
  colnames(pt_df)[1:2] <- c('Cluster','Treatment')
  pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(seuratObj$treatment))
  
  alluvial_plot_orig_ident = ggplot(pt_df,aes(x = Treatment, 
                                              stratum = Cluster, 
                                              alluvium = Cluster,
                                              y = Freq,
                                              fill = Cluster)) +
    geom_stratum(alpha = 1) +
    geom_flow(alpha = 0.01) +
    theme_bw(base_size = 15) +
    geom_flow(stat = "alluvium", lode.guidance = "forward") +
    scale_fill_manual(values=seuratObj@misc$cols_seurat[rownames(pt)])
  
  alluvial_plot_orig_ident  
}

rp <- relative_plots(glioma_subsets[[1]])
for(i in 2:length(glioma_subsets)) {
  rp <- rp + relative_plots(glioma_subsets[[i]])
}
rp
ggsave('Figures/cluster_abs_frequency.png', 
       height = 10, width = 12)
```

### Per group, absolute Barplots

```{r, fig.height=7, fig.width=10, message=FALSE}
library(ggplot2)
library(ggalluvial)

glioma_subsets <- list(
  tumor=glioma_tumor_subset, 
  lymphatic=glioma_lymphatic_subset, 
  parenchymal=glioma_parenchymal_subset, 
  myeloid=glioma_myeloid_subset)

n_cells <- rowSums(table(glioma_integrated$treatment, droplevels(glioma_integrated$seurat_clusters)))

absolute_plots <- function(seuratObj, n_cells) {
  pt <- table(seuratObj$treatment, droplevels(seuratObj$seurat_clusters))
  pt <- t(as.matrix(pt)/n_cells)
  pt_df <- as.data.frame(pt)
  colnames(pt_df)[1:2] <- c('Cluster','Treatment')
  pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(seuratObj$treatment))
  
  alluvial_plot_orig_ident = ggplot(pt_df,aes(x = Treatment, 
                                              stratum = Cluster, 
                                              alluvium = Cluster,
                                              y = Freq,
                                              fill = Cluster)) +
    geom_stratum(alpha = 1) +
    geom_flow(alpha = 0.01) +
    theme_bw(base_size = 15) +
    geom_flow(stat = "alluvium", lode.guidance = "forward") +
    scale_fill_manual(values=seuratObj@misc$cols_seurat[rownames(pt)])
  
  alluvial_plot_orig_ident  
}

rp <- absolute_plots(glioma_subsets[[1]], n_cells)
for(i in 2:length(glioma_subsets)) {
  rp <- rp + absolute_plots(glioma_subsets[[i]], n_cells)
}
rp
ggsave('Figures/cluster_rel_frequency.png', height = 10, width = 12)
```

```{r}
glioma_subsets <- list(
  tumor=glioma_tumor_subset, 
  lymphatic=glioma_lymphatic_subset, 
  parenchymal=glioma_parenchymal_subset, 
  myeloid=glioma_myeloid_subset)

n_cells <- rowSums(table(glioma_integrated$treatment, droplevels(glioma_integrated$seurat_clusters)))

make_prop_table <- function(seuratObj, filename) {
  
  pt <- table(seuratObj$treatment, droplevels(seuratObj$seurat_clusters))
  
  cell_df <- as.data.frame(pt)
  colnames(cell_df)[3] <- 'Cells'
  freq_df <- as.data.frame(as.matrix(pt)/rowSums(pt))
  colnames(freq_df)[3] <- 'Rel. Freq.'
  abs <- as.matrix(pt)/n_cells
  abs_df <- as.data.frame(abs)
  colnames(abs_df)[3] <- 'Abs. Freq.'
  
  csv_df <- merge(cell_df, freq_df)
  csv_df <- merge(csv_df, abs_df)

  colnames(csv_df)[1:2] <- c('Treatment','Cluster')
  return(csv_df)

}

openxlsx::write.xlsx(lapply(glioma_subsets, make_prop_table), 
                   file = 'Tables/proportions_cells_treatments.xlsx')
```


### absolute, per group Barplot

```{r}
tumor_group <- c('5','8','9','10')
lymphatic_group <- c('1','4','6','14','16','19','21','22','26','24','27')
parenchymal_group <- c('11','13','17','23','25','28')
myeloid <- setdiff(as.character(sort(unique(glioma_integrated$seurat_clusters))), 
                   c(tumor_group, lymphatic_group, parenchymal_group))

# assign the groups within the main object
cell_group <- rep('tumor', ncol(glioma_integrated))
cell_group[glioma_integrated$seurat_clusters %in% lymphatic_group] <- 'lymphatic'
cell_group[glioma_integrated$seurat_clusters %in% parenchymal_group] <- 'parenchymal'
cell_group[glioma_integrated$seurat_clusters %in% myeloid] <- 'myeloid'
glioma_integrated$cell_group <- cell_group

pt <- table(glioma_integrated$treatment, glioma_integrated$cell_group)
pt <- t(as.matrix(pt)/rowSums(pt))
pt_df <- as.data.frame(pt)
colnames(pt_df)[1:2] <- c('Cluster','Treatment')
pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(glioma_integrated$treatment))

alluvial_plot_orig_ident = ggplot(pt_df,aes(x = Treatment, 
                                            stratum = Cluster, 
                                            alluvium = Cluster,
                                            y = Freq,
                                            fill = Cluster)) +
  geom_stratum(alpha = 1) +
  geom_flow(alpha = 0.01) +
  theme_bw(base_size = 15) +
  geom_flow(stat = "alluvium", lode.guidance = "forward") +
  scale_fill_manual(values=glioma_integrated@misc$cols_groups[rownames(pt)])

alluvial_plot_orig_ident
ggsave('Figures/groups_abs_frequency.png', 
       height = 5, width = 7)
```

# CellCycle

```{r}
s.genes <- intersect(stringr::str_to_title(cc.genes$s.genes), rownames(glioma_integrated_RNA))
g2m.genes <- intersect(stringr::str_to_title(cc.genes$g2m.genes), rownames(glioma_integrated_RNA))
glioma_integrated_RNA <- CellCycleScoring(glioma_integrated_RNA, 
                                          s.features = s.genes, g2m.features = g2m.genes)
```

```{r}
cs1 <- FeaturePlot(glioma_integrated_RNA, features = 'S.Score', max.cutoff = 10, min.cutoff = 0)
cs2 <- FeaturePlot(glioma_integrated_RNA, features = 'G2M.Score', max.cutoff = 10, min.cutoff = 0)
cs3 <- DimPlot(glioma_integrated_RNA, group.by = 'Phase')
```

```{r, fig.height=5, fig.width=10}
cs1 + cs2 + plot_layout(ncol=2, guides = 'collect')
ggsave('Figures/umap_filtered_cc_scores.png', 
       height = 4, width = 8)
```

```{r}
cs3
ggsave('Figures/umap_filtered_cc_phase.png', 
       height = 4, width = 4.5)
```

```{r, fig.height=8, fig.width=8}
DimPlot(glioma_integrated_RNA, group.by = 'Phase', split.by = 'treatment', ncol=2)
ggsave('Figures/umap_filtered_cc_phase_splitTreatment.png', 
       height = 8, width = 9)
```

```{r, fig.height=7, fig.width=10}
library(ggplot2)
library(ggalluvial)

#--- tumor genotype
pt <- with(subset(glioma_integrated@meta.data, tumor_assignment=='tumor'), table(treatment, Phase))
pt <- t(as.matrix(pt)/rowSums(pt))
pt_df <- as.data.frame(pt)
colnames(pt_df)[1:2] <- c('Phase','Treatment')
pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(glioma_integrated$treatment))

alluvial_plot_tumor_cells = ggplot(pt_df,aes(x = Treatment, stratum = Phase, alluvium = Phase,
              y = Freq,
              fill = Phase)) +
  geom_stratum(alpha = 1) +
  geom_flow(alpha = 0.01) +
  theme_bw(base_size = 15) +
  geom_flow(stat = "alluvium", lode.guidance = "forward") + ggtitle('tumor')

#--- normal genotype
pt <- with(subset(glioma_integrated@meta.data, tumor_assignment=='normal'), table(treatment, Phase))
pt <- t(as.matrix(pt)/rowSums(pt))
pt_df <- as.data.frame(pt)
colnames(pt_df)[1:2] <- c('Phase','Treatment')
pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(glioma_integrated$treatment))

alluvial_plot_normal_cells = ggplot(pt_df,aes(x = Treatment, stratum = Phase, alluvium = Phase,
              y = Freq,
              fill = Phase)) +
  geom_stratum(alpha = 1) +
  geom_flow(alpha = 0.01) +
  theme_bw(base_size = 15) +
  geom_flow(stat = "alluvium", lode.guidance = "forward") + ggtitle('normal')

#--- all genotype
pt <- with(glioma_integrated@meta.data, table(treatment, Phase))
pt <- t(as.matrix(pt)/rowSums(pt))
pt_df <- as.data.frame(pt)
colnames(pt_df)[1:2] <- c('Phase','Treatment')
pt_df$Treatment <- factor(pt_df$Treatment, levels=levels(glioma_integrated$treatment))

alluvial_plot_all_cells = ggplot(pt_df,aes(x = Treatment, stratum = Phase, alluvium = Phase,
              y = Freq,
              fill = Phase)) +
  geom_stratum(alpha = 1) +
  geom_flow(alpha = 0.01) +
  theme_bw(base_size = 15) +
  geom_flow(stat = "alluvium", lode.guidance = "forward") + ggtitle('all')

alluvial_plot_tumor_cells + alluvial_plot_normal_cells + alluvial_plot_all_cells + 
  plot_layout(guides = 'collect')
ggsave('Figures/cc_phase_abs_frequency.png',
       height = 5, width = 14)
```


# Gene lists

## Classification genes 

```{r}
gene_list <- c("Gfap","Itgam","Itgax","Tmem119","Cx3cr1","Scin","P2ry12","Sall1","Slc2a5","Ly6c1","Ccr2","Ptprc","Itga2","Cd3g","Cd4","Cd8a","Il2ra","Adgre1","Cd68","Foxp3","Pdcd1","Cd274","Pdcd1lg2","Lag3")
```

```{r}
# cluster_groups
# glioma_integrated@misc$cols_groups
```

```{r}
view_out <- viewGeneList(gene_list, 
                         groups_order = c('lymphatic','myeloid','parenchymal','tumor'),
                         ncol = 3)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels), fig.width=12}
view_out$fp
ggsave('Figures/features_classification.png', 
       height = 1+length(view_out$avg_heat$tree_row$labels), 
       width = 12)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels)/3, fig.width=10}
as.ggplot(view_out$avg_heat_by_group)
ggsave('Figures/heatmap_classification.png',
       height = 1+length(view_out$avg_heat$tree_row$labels)/3,
       width = 10
       )
```

## TAMs_M1

```{r}
gene_list <- c(
'Fcgr3a', # CD16  TAMs_M1
'Cd86', # TAMs_M1
'Il12a', # IL-12 TAMs_M1
'Tnf', # TNF-α TAMs_M1
'Ifng', # IFN-γ TAMs_M1
'Nos2')
```

```{r}
view_out <- viewGeneList(gene_list, ncol = 3)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels), fig.width=12}
view_out$fp
ggsave('Figures/features_TAMs_M1.png',
       height = 1+length(view_out$avg_heat$tree_row$labels),
       width = 12)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels)/3, fig.width=10}
as.ggplot(view_out$avg_heat_by_group)
ggsave('Figures/heatmap_TAMs_M1.png',
       height = 1+length(view_out$avg_heat$tree_row$labels)/3,
       width = 10)
```

## TAMs_M2

```{r}
gene_list <- c('Mrc1', # CD206(q)   TAMs_M2
'Il4', # IL-4  TAMs_M2
'Cxcl8', # IL-8  TAMs_M2
'Il10', # IL-10 TAMs_M2
'Il13', # IL-13 TAMs_M2
'Tgfb1', # TGF-β TAMs_M2
'Arg1') # Arg-1 TAMs_M2
```

```{r}
view_out <- viewGeneList(gene_list, ncol = 3)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels), fig.width=12}
view_out$fp
ggsave('Figures/features_TAMs_M2.png',
       height=1+length(view_out$avg_heat$tree_row$labels), 
       width=12
       )
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels)/3, fig.width=10}
as.ggplot(view_out$avg_heat_by_group)
ggsave('Figures/heatmap_TAMs_M2.png',
       height=1+length(view_out$avg_heat$tree_row$labels)/3, 
       width=10
       )
```

## Tumor response

```{r}
gene_list <- intersect(c('Trem2','Tspo','Fap','Cxcr4'), rownames(glioma_integrated_RNA))
```

```{r}
view_out <- viewGeneList(gene_list, ncol = 3)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels), fig.width=12}
view_out$fp
ggsave('Figures/features_tumor.png', 
       height=1+length(view_out$avg_heat$tree_row$labels), 
       width=12)
```

```{r, fig.height=1+length(view_out$avg_heat$tree_row$labels)/3, fig.width=10}
as.ggplot(view_out$avg_heat_by_group)
ggsave('Figures/heatmap_tumor.png',
       height=1+length(view_out$avg_heat$tree_row$labels)/3, 
       width=10
       )
```





